Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
more editing
small change to section title
removed the pseudocode for grammar induction
shortened bibliography

We define a prior distribution over DSLs which penalizes the sizes of the $\lambda$-calculus expressions in the DSL, and put a Dirichlet prior over the weight vector:
\begin{align*}
  \probability[\mathcal{D}]&\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\\
  P(\theta|\mathcal{D})& = \text{Dir}(\theta|\alpha)
\end{align*}
where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
$\lambda$ is a hyperparameter that acts as a regularizer on the size of the DSL,
and $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar,
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
a DSL in \system may be context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Supplement we derive a tractable MAP estimator for $\theta$.
\begin{algorithm}[tb]
   \caption{DSL Induction Algorithm}
   \label{grammarInductionAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
     \WHILE {true}
     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
     \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
     \STATE $\mathcal{D}\gets\mathcal{D}'$
     \ENDWHILE
   \end{algorithmic}
\end{algorithm}

@@

\subsection{Implementing the \system}

@@

We compare  with four baselines on  held-out tasks:
\\\noindent \textbf{Ours--NN}, which lesions the recognition model.
\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (see Sec.~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (see Sec.~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{PCFG}, which lesions the recognition model, learns $\theta$, and fixes $\mathcal{D}$.
This is equivalent to our algorithm with $q(x) = \theta$ and $\lambda = \infty$,
and is like learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which does no learning and just enumerates a frontier.
This is equivalent to the first wake cycle of \system.

@@

 Our work is an instance of
 Bayesian Program
 Learning (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}). Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),
 and our contribution here is a general scheme for getting BPL off the ground with less hand-engineering of the DSL.

@@

\pagebreak

>>>>

We define a prior distribution over DSLs which penalizes the sizes of the $\lambda$-calculus expressions in the DSL, and put a Dirichlet prior over the weight vector:
\begin{align*}
  \probability[\mathcal{D}]&\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\\
  P(\theta|\mathcal{D})& = \text{Dir}(\theta|\alpha)
\end{align*}
where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
$\lambda$ is a hyperparameter that acts as a regularizer on the size of the DSL,
and $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar,
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
a DSL in \system may be context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Supplement we derive a tractable MAP estimator for $\theta$.
%% \begin{algorithm}[tb]
%%    \caption{DSL Induction Algorithm}
%%    \label{grammarInductionAlgorithm}
%%    \begin{algorithmic}
%%      \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
%%      \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
%%      \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
%%      \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
%%      \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
%%      \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
%%      \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
%%      \WHILE {true}
%%      \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
%%      \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
%%      \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
%%      \STATE $\mathcal{D}\gets\mathcal{D}'$
%%      \ENDWHILE
%%    \end{algorithmic}
%% \end{algorithm}

@@

\subsection{Implementing \system}

@@

We compare  with four baselines on  held-out tasks:
\\\noindent \textbf{Ours--NN}, which lesions the recognition model.
\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (see Sec.~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (see Sec.~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{PCFG}, which lesions the recognition model, learns $\theta$, and fixes $\mathcal{D}$.
This is equivalent to our algorithm with $q(x) = \theta$ and $\lambda = \infty$,
and is like learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which does no learning and just enumerates a frontier.
This is equivalent to our first wake cycle.

@@

 Our work is an instance of
 Bayesian Program
 Learning (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}). Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),
 and our contribution here is a general way of doing BPL with less hand-engineering of the DSL.

@@
